---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zoomclass

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

> **A high-level data analysis R package for Zoom's participants report `.csv` and Zoom's chat `.txt` file.**

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("Lightbridge-KS/zoomclass")
```

Functions in [{zoomclass}](https://github.com/Lightbridge-KS/zoomclass) package are divided in 2 categories which aims to analyse input from (mainly) Zoom's participants report `.csv` and Zoom's chat `.txt` file.

```{r message=FALSE, warning=FALSE}
library(zoomclass)
```

# Participants Report

In this category, High-level functions are design to analyse Zoom's participants report (`.csv` file), especially for checking student's attendance in Zoom classroom.

## Zoom's Participant Report

The term I used "Zoom's participant report", basically, is a CSV log file that host can downloaded after Zoom meeting has ended. It contains the following columns:

-   **Name (Original Name):** show participant's *current name* in Zoom meeting with optional *original name* in the last balanced parenthesis.
-   **User Email:** an email that participant signed in Zoom account
-   **Join Time:** join time of each participants by each individual sessions
-   **Leave Time:** leave time of each participants by each individual sessions
-   **Duration (Minutes):** computed by `Leave Time` - `Join Time` of each session
-   **Guest:** "Yes" means that person is a participants; "No" means he/she is a host
-   **Recording Consent:** "Y" means participant allowed recording in zoom

This package comes with an example data of Zoom's participant report in `.csv` which I will use for demo in this tutorial.

```{r zoomclass_example}
# List all example data in `zoomclass` package
zoomclass_example()

# Get a path to specific example data
path_heroes <- zoomclass_example("participants_heroes.csv")
path_heroes
```

You can see all example files in file explorer by executing this command.

```{r showfile-example}
if (interactive() && requireNamespace("fs")) {

  fs::file_show(system.file("extdata", package = "zoomclass"))
  
}
```

## Read Participants

`read_participants()` read the data in using `readr::read_csv()` and then performs the followings steps:

1.  Read Zoom's participant report into a tibble
2.  Clean column names, so that we can manipulate more easily in `R`
3.  Apply appropriate time formatting (POSIXct) for relevant columns
4.  Extract participant's current (displayed) name and original name into `Name` and `Name_Original` columns, respectively.

The output is a tibble with `zoom_participants` subclass. If metadata regarding Zoom meeting was found, it will be assign to a `meeting_overview` attribute.

```{r pp_heroes}
pp_heroes <- read_participants(path_heroes)

str(pp_heroes)
```

As you can see, almost all column names were cleaned with no spaces; `Join_Time` and `Leave_Time` were formatted as POSIXct object, correctly. Moreover, "015_Loki (Loki Laufeyson)" in original `Name (Original Name)` column is extracted into "015_Loki" in `Name` and "Loki Laufeyson" in `Name_Original`, respectively.

The output tibble `pp_heroes` contain sessions information of each participants. If we count names of each participants, we can see that some rows is duplicated.

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

```{r}

pp_heroes %>% 
  count(`Name (Original Name)`, sort = TRUE) %>% 
  head()
```

For example, "Power Girl" has 2 rows because she had 2 sessions in Zoom (join and leave 2 times).

```{r}
pp_heroes %>% 
  filter(Name == "003_Power Girl") %>% 
  select(Name, Email, Join_Time, Leave_Time, Duration_Minutes)
```

But why `Join_Time` and `Leave_Time` is overlapped in these 2 sessions? This might be the case that she's joined Zoom with 2 devices at the same time. As I will show you next, The upcoming functions in this package can detect this scenario.

## Classroom in Zoom


Now that we have read Zoom's participants data in, It's time to analyse participants in a context of a classroom. From now on, I'll call a "participant" as "student".

A typical academic classroom usually has an explicit start and end time. If students arrive to class later than certain cutoff time point, teacher can mark them as late.

In the next family of functions: `class_*`, you must provide `class_start` and `class_end` arguments by which they will used to compute 4 time intervals in the corresponding columns:

-   `Before_Class`: represent time interval that student spent in Zoom before class started
-   `During_Class`: represent time interval that student spent in Zoom during class
-   `After_Class`: represent time interval that student spent in Zoom after class ended
-   `Total_Time`: the sum of time `Before_Class` + `During_Class` + `After_Class`

All of theses time periods are Period object from `{lubridate}` package. (It can be converted to hours, minutes, or secound as well.)

Furthermore, `class_*` functions can tell you whether each students joined Zoom with multiple devices at the same time period.
The use case of this might be when live-exam is conducted in Zoom, and you want to check that each students joined with 1 device only (no cheating).

Next, 3 **`class_*`** functions will be introduced:

1. **`class_session()`** summarizes time information about individual sessions of each students (If student has multiple sessions, output will show ≥ 1 rows per that student)
2. **`class_students()`** summarizes time information of each students (1 row per student)
3. **`class_students()`** summarizes time information of each student's ID extracted from student name (1 row per student's ID)


The first argument of these functions are `zoom_participants` tibble as created by `read_participants()`.



### Class Session

Suppose that our Zoom classroom was started at 10:00 and ended at 12:00 (24 hours clock time), I will call `class_session()` as follows:


```{r class_session}
pp_heroes_session <- 
  class_session(pp_heroes, 
                class_start = "10:00",
                class_end = "12:00"
                )

pp_heroes_session
```

As previously stated, classroom-related time intervals were computed i.e., `Before_Class`, `During_Class`, and `After_Class`.

Let's see who spend time during class all the time

```{r}
pp_heroes_session %>% 
  filter(During_Class == lubridate::hours(2)) %>% 
  select(Name, ends_with("Class"))
```

Or, may be Loki could cast illusion to attend Zoom till the end :-) 


Now, let's see who attend Zoom using multiple device at the same time by filter `Multi_Device = TRUE`. 



```{r}
pp_heroes_session %>% 
  filter(Multi_Device  == TRUE) %>% 
  select(Name, Session, Join_Time, Leave_Time, Multi_Device)
```
`Session` column displays session number of each student as ranked by `Join_Time`. Let's see who joined Zoom more than 1 time.

As you can see, "Power Girl" and "Swarm" has session number = 2; however, be aware that these two joined Zoom only one time using 2 devices.


```{r}
pp_heroes_session %>% 
  filter(Session  > 1) %>% 
  select(Name, Session, Join_Time, Leave_Time, Multi_Device)
```

### Class Student (TO DO)


### Class Student ID (TO DO)


## Zoom Chat

The final functions of this package is design to parse program Zoom [chat file](https://support.zoom.us/hc/en-us/articles/115004792763-Saving-in-meeting-chat) from `.txt` file to a tibble, just execute the followings:

```{r example, eval=FALSE}
read_zoom_chat("path/to/zoom_chat.txt")
```
