---
title: "Introduction to zoomclass"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to zoomclass}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Functions in [{zoomclass}](https://github.com/Lightbridge-KS/zoomclass) package are divided in 2 categories which aims to analyse input from (mainly) Zoom's participants report `.csv` and Zoom's chat `.txt` file.

```{r message=FALSE, warning=FALSE}
library(zoomclass)
```

# Participants Report

In this category, High-level functions are design to analyse Zoom's participants report (`.csv` file), especially for checking student's attendance in Zoom classroom.

## Zoom's Participant Report

The term I used "Zoom's participant report", basically, is a CSV log file that host can downloaded after Zoom meeting has ended. It contains the following columns:

-   **Name (Original Name):** show participant's *current name* in Zoom meeting with optional *original name* in the last balanced parenthesis.
-   **User Email:** an email that participant signed in Zoom account
-   **Join Time:** join time of each participants by each individual sessions
-   **Leave Time:** leave time of each participants by each individual sessions
-   **Duration (Minutes):** computed by `Leave Time` - `Join Time` of each session
-   **Guest:** "Yes" means that person is a participants; "No" means he/she is a host
-   **Recording Consent:** "Y" means participant allowed recording in zoom

This package comes with an example data of Zoom's participant report in `.csv` which I will use for demo in this tutorial.

```{r zoomclass_example}
# List all example data in `zoomclass` package
zoomclass_example()

# Get a path to specific example data
path_heroes <- zoomclass_example("participants_heroes.csv")
path_heroes
```


Path to all example files can be obtained by the following command.

```{r showfile-example, include=FALSE}
# if (interactive() && requireNamespace("fs")) {
#   fs::file_show(system.file("extdata", package = "zoomclass"))
# }
```

```{r extdata_path}
system.file("extdata", package = "zoomclass")
```


## Read Participants

`read_participants()` read the data in using `readr::read_csv()` and then performs the followings steps:

1.  Read Zoom's participant report into a tibble
2.  Clean column names, so that we can manipulate more easily in `R`
3.  Apply appropriate time formatting (POSIXct) for relevant columns
4.  Extract participant's current (displayed) name and original name into `Name` and `Name_Original` columns, respectively.

The output is a tibble with `zoom_participants` subclass. If metadata regarding Zoom meeting was found, it will be assigned to a `meeting_overview` attribute.

```{r pp_heroes}
pp_heroes <- read_participants(path_heroes)

str(pp_heroes)
```

As you can see, almost all column names were cleaned with no spaces; `Join_Time` and `Leave_Time` were formatted as POSIXct object, correctly. Moreover, "015_Loki (Loki Laufeyson)" in original `Name (Original Name)` column is extracted into "015_Loki" in `Name` and "Loki Laufeyson" in `Name_Original`, respectively.

The output tibble `pp_heroes` contain sessions information of each participants. If we count names of each participants, we can see that some rows is duplicated.

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

```{r}
pp_heroes %>% 
  count(`Name (Original Name)`, sort = TRUE) %>% 
  head()
```

For example, "Power Girl" has 2 rows because she had 2 sessions in Zoom (join and leave 2 times).

```{r}
pp_heroes %>% 
  filter(Name == "003_Power Girl") %>% 
  select(Name, Email, Join_Time, Leave_Time, Duration_Minutes)
```

But why `Join_Time` and `Leave_Time` is overlapped in these 2 sessions? This might be the case that she's joined Zoom with 2 devices at the same time. As I will show you next, The upcoming functions in this package can detect this scenario.

## Classroom in Zoom

Now that we have read Zoom's participants data in, It's time to analyse participants in a context of a classroom. From now on, I'll call a "participant" as "student".

Next, 3 **`class_*`** functions will be introduced:

1.  **`class_session()`** summarizes time information about individual sessions of each students (If student has multiple sessions, output will show ≥ 1 rows per that student)
2.  **`class_students()`** summarizes time information of each students (1 row per student)
3.  **`class_studentsID()`** summarizes time information of each student's ID extracted from student name (1 row per student's ID)

The first argument of these functions receives input from `zoom_participants` tibble as created by `read_participants()`.

A typical academic classroom usually has an explicit start and end time. If students arrive to class later than certain cutoff time point, teacher can mark them as late.

In the `class_*` functions, you must provide `class_start` and `class_end` arguments by which they will used to compute 4 time intervals that student spent before, during, and after Zoom class:

-   `Before_Class`: represent time interval that student spent in Zoom before class started
-   `During_Class`: represent time interval that student spent in Zoom during class
-   `After_Class`: represent time interval that student spent in Zoom after class ended
-   `Total_Time`: the sum of time `Before_Class` + `During_Class` + `After_Class`

For `class_students()` and `class_studentsID()`, you can compute late time period of each student by providing `late_cutoff` argument.

All of theses time intervals are Period object from `{lubridate}` package. (It can be converted to hours, minutes, or secound as well.)

Furthermore, `class_*` functions can tell you whether each students joined Zoom with multiple devices at the same time period. The use case of this might be when live-exam is conducted in Zoom, and you want to check that each students joined with 1 device only (no cheating).



### Class Session

Suppose that our Zoom classroom was started at 10:00 and ended at 12:00, I will call `class_session()` as follows:

(Input `class_start` and `class_end` as *24-hours clock time* with no AM or PM)

```{r class_session}
pp_heroes_session <- 
  class_session(pp_heroes, 
                class_start = "10:00", # Official class started at 10:00 AM
                class_end = "12:00" # Official class ended at 12:00 PM
                )

pp_heroes_session
```

As previously stated, classroom-related time intervals were computed.


```{r}
pp_heroes_session %>% 
  select(Name, ends_with("Class"), Total_Time) %>% 
  head()
```


Let's see who spend time during class all the time

```{r}
pp_heroes_session %>% 
  filter(During_Class == lubridate::hours(2)) %>% 
  select(Name, ends_with("Class"))
```

Or, may be Loki could cast illusion to attend Zoom till the end :-)

Now, let's see who attend Zoom using multiple device at the same time by filter `Multi_Device = TRUE`.

```{r}
pp_heroes_session %>% 
  filter(Multi_Device  == TRUE) %>% 
  select(Name, Session, Join_Time, Leave_Time, Multi_Device)
```

`Session` column displays session number of each student as ranked by `Join_Time`. Let's see who joined Zoom more than 1 time.

As you can see, "Power Girl" and "Swarm" has session number = 2; however, be aware that these two joined Zoom only one time using 2 devices.

```{r}
pp_heroes_session %>% 
  filter(Session  > 1) %>% 
  select(Name, Session, Join_Time, Leave_Time, Multi_Device)
```

### Class Student

**`class_students()`** summarizes time information grouped by each students (grouping variables is `Name (Original Name)` and `Email`).
The result will be a tibble with one row per student. 

Most columns are the same as output from `class_session()`, except for the followings:

-   `Session_Count`: show session counts of each student (how many times each student join and leave class)
-   `First_Join_Time`: if student has multiple sessions, this would choose only the earliest join time.
-   `Last_Leave_Time`: if student has multiple sessions, this would choose only the latest leave time.

New optional argument is `late_cutoff` from which student will be considered late if first joined time is later than this cutoff. The late time period will be shown in `Late_Time` column.

```{r pp_heroes_students}
pp_heroes_students <- 
  class_students(pp_heroes, 
               class_start = "10:00",
               class_end = "12:00",
               late_cutoff = "10:15" # If student joined later than 10:15 will considered late
              )

head(pp_heroes_students)
```


Let's see who joined class late > 10 minutes (later than 10:25).

```{r}
pp_heroes_students %>% 
  filter(Late_Time > lubridate::minutes(10)) %>% 
  select(Name, First_Join_Time, Late_Time)
```



### Class Student ID

Supposed that after you've checked class attendance of each student, perhaps you want to merge this data into a database by some key columns which is usually student's ID. 

First, you informed students to put student's ID in their names, for example: "001_Megus".

**`class_studentsID()`** will help you summarizes time information grouped by each student's ID. Then, you can used these IDs to merge into a database.


The internal processes are that `class_studentsID()` extracts student's ID from `Name (Original Name)` column using regular expression as provided by `id_regex`. Then, time information will be summarized per student's ID (grouping variable is `ID`), and the rest of output columns is similar to `class_students()`. Finally, the result will be a tibble with one row per `ID`. 



```{r pp_heroes_studentsID}
pp_heroes_studentsID <- 
  class_studentsID(pp_heroes, 
                   id_regex = "\\d+", # Extract digits from student name as student's ID
                   class_start = "10:00",
                   class_end = "12:00",
                   late_cutoff = "10:15" # If student joined later than 10:15 will considered late
                   )

head(pp_heroes_studentsID)
```

(If the same `ID` is founded in multiple names (e.g. "001_Magus", "001_magus"), the `Name` column will contain all combinations of names for that particular `ID`.)


### Merge Data

Here, I will give an example of merging student data to a database.

This package provides `heroes_students` data frame that contain names and ID of student who enrolled the course in this semester.

```{r}
head(heroes_students)
```

You can join `heroes_students` with `pp_heroes_studentsID` using `dplyr::*_join` functions by `ID`.



#### Check Students NOT in Zoom Class

To check whether students joined Zoom class room or not, it can be obtained by using a filtering join function: `dplyr::anti_join()`. 


These students below are in the `heroes_students` data frame, but not in `pp_heroes_studentsID`, which means that they didn't join Zoom classroom.

```{r}
heroes_students %>% 
  anti_join(pp_heroes_studentsID, by = "ID")
```

#### Check Non-Student in Zoom Class

Likewise, You can also check participants who joined Zoom classroom but not in the `heroes_students` data frame (non-students). Again, `dplyr::anti_join()` can be used.

```{r}
pp_heroes_studentsID %>% 
  anti_join(heroes_students, by = "ID")
```

#### Merge All 

`dplyr::full_join()` is a mutating join function that merge 2 data frame without any rows lost from either one. 

In this example, I fully joined 2 data frame `pp_heroes_studentsID` and `heroes_students` by `ID`. Suffix "_from_ID" and "_from_Zoom" represents unmatched rows from the list of students (`heroes_students`) and Zoom classroom, respectively.


```{r heroes_students_joined}
# Merge to Zoom data by `ID`
heroes_students_joined <- heroes_students %>% 
  full_join(pp_heroes_studentsID, by = "ID", suffix = c("_from_ID", "_from_Zoom")) %>% 
  relocate(ID, starts_with("Name"))

head(heroes_students_joined)
```

Student who didn't joined Zoom classroom will have `NA` presented in the `Name_from_Zoom` column, whereas participants who joined Zoom classroom but not in the list of students will have `NA` presented in the `Name_from_ID` column.


```{r}
heroes_students_joined %>% 
  filter(if_any(starts_with("Name"), is.na)) %>% 
  select(ID, starts_with("Name"), Comment)
```



## Zoom Chat

The final functions of this package is design to parse program Zoom [chat file](https://support.zoom.us/hc/en-us/articles/115004792763-Saving-in-meeting-chat) from `.txt` file to a tibble.
j

```{r path}
# Path to example Zoom Chat file
path <- zoomclass_example("zoom-chat-1.txt")
```


Raw `zoom-chat-1.txt` looks like this.

```{r}
readLines(path) %>% 
  glue::as_glue()
```


Use `read_zoom_chat()` to read it into a data frame.

```{r zoomchat}
# Read from Text to a Data Frame
read_zoom_chat(path)
```

---
Last updated: `r Sys.Date()`
